# What is **SWIM**? {#Sec:Intro}

```{r wrap-hook, echo = FALSE}
library(knitr)
hook_error = knit_hooks$get('error')
knit_hooks$set(error = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_error(x, options)
})
```



## Sensitivity testing and scenario weights

The purpose of **SWIM** is to enable sensitivity analysis of models implemented in a Monte Carlo simulation framework, by distorting (*stressing*) some of the models' components and monitoring the resulting impact on quantities of interest.
To clarify this idea and explain how **SWIM** works, we first define the terms used. By a *model*, we mean a set of $n$ (typically simulated) realisations from a vector of random variables $(X_1,\dots,X_d)$, along with *scenario weights* $W$ assigned to individual realisations, as shown in the table below. Hence each of of the columns 1 to $d$ corresponds to a random variable, called a *model component*, while each row corresponds to a *scenario*, that is, a state of the world.

| $X_1$   	| $X_2$ 	    | $\dots$ 	| $X_d$ 	   | $W$ 	|
|:--------:	|:----:      |:------:  |:-----:    |:-------------:	|
| $x_{11}$ 	|$x_{21}$    | $\dots$  | $x_{d1}$  | $w_1$            	|
| $x_{12}$ 	|$x_{22}$    | $\dots$  | $x_{d2}$  | $w_2$            	|
| $\vdots$  |$\vdots$    | $\ddots$ | $\vdots$  | $\vdots$           |
| $x_{1n}$  | $x_{2n}$   | $\dots$  | $x_{dn}$  | $w_n$             |
Table: (\#tab:SWIMframework) Illustration of the **SWIM** framework, that is the baseline model, the stressed model and the scenario weights.  

Each scenario has a *scenario weight*, shown in the last column, such that, scenario $i$ has probability $\frac{w_i}{n}$ of occurring. Scenario weights are always greater or equal than zero and have an average of 1. When all scenario weights are equal to 1, such that the probability of each scenario is $\frac 1 n$ (the standard Monte Carlo framework), we call the model a *baseline model* -- consequently weights of a baseline model will never be explicitly mentioned. When scenario weights are not identically equal to 1, such that some scenarios are more heavily weighted than others, we say that we have a *stressed model*. 

The scenario weights make the joint distribution of model components under the stressed model  different, compared to the baseline model. For example, under the baseline model, the expected value of $X_1$ and the cumulative distribution function of $X_1$, at threshold $t$, are respectively given by:

$$
E(X_1)=\frac 1  n \sum_{i=1}^nx_{1i},\quad F_{X_1}(t)= P(X_1\leq t)=\frac 1 n \sum_{i=1}^n \mathbf 1 _{x_{1i}\leq t},
$$

where $\mathbf 1 _{x_{1i}\leq t}=1$ if $x_{1i}\leq t$ and $0$ otherwise. For a stressed model with scenario weights $W$, the expected value $E^W$ and cumulative distribution function $F^W$ become:

$$
E^W(X_1)=\frac 1  n \sum_{i=1}^n w_i x_{1i},\quad F_{X_1}^W(t)=P^W(X_1\leq t)=\frac 1 n \sum_{i=1}^n w_i \mathbf 1 _{x_{1i}\leq t}.
$$

Similar expressions can be derived for more involved quantities, such as higher (joint) moments and quantiles. 

The logic of stressing a model with **SWIM** then proceeds as follows. An analyst or modeller is  supplied with a baseline model, in the form of a matrix of equiprobable simulated scenarios of model components. The modeller wants to investigate the impact of a change in the distribution of, say, $X_1$. To this effect, she chooses a *stress* on the distribution of $X_i$, for example requiring that $E^W(X_1)=m$; we then say that she is *stressing* $X_1$ and, by extension, the model. Subsequently, **SWIM** calculates the scenario weights such that the stress is fulfilled and the distortion to the baseline model induced by the stress is as small as possible; specifically the Kullback-Leibler divergence (or relative entropy) between the baseline and stressed models is minimised. (See Section \@ref(Rfunctions) for more detail on the different types of possible stresses and the corresponding optimisation problems). Once scenario weights are obtained, they can be used to determine the stressed distribution of any model component or function of model components. For example, for scenario weights $W$ obtained through a stress on $X_1$, we may calculate

$$
E^W(X_2)=\frac 1  n\sum_{i=1}^n w_i x_{2i},\quad E^W(X_1^2+X_2^2)=\frac 1  n \sum_{i=1}^n w_i \left(x_{1i}^2+ x_{2i}^2 \right).
$$

Through this process, the modeller can monitor the impact of the stress on $X_1$ on any other random variable of interest. It is notable that this approach does not necessitate generating new simulations from a stochastic model. As the **SWIM** approach requires a single set of simulated scenarios (the baseline model) it offers a clear computational benefit. 

## An introductory example
Here, through an example, we illustrate the basic concepts and usage of **SWIM** for sensitivity analysis. More advanced usage of **SWIM** and options for constructing stresses are demonstrated in Sections \@ref(Sec:Scope) and \@ref(Sec:CreditModel).
We consider a simple portfolio model, with the portfolio loss defined by $Y=Z_1+Z_2+Z_3$. The random variables $Z_1,Z_2,Z_3$ represent normally distributed losses, with $Z_1\sim N(100,40^2)$, $Z_2\sim Z_3\sim N(100,20^2)$. $Z_1$ and $Z_2$ are correlated, while $Z_3$ is independent of $(Z_1,Z_2)$.  Our purpose in this example is to investigate how a stress on the loss $Z_1$ impacts on the overall portfolio loss $Y$.
First we derive simulated data from the random vector $(Z_1,Z_2,Z_3,Y)$, forming our baseline model.

``` {r, example1_sim_data, include = TRUE}
set.seed(0)
# number of simulated scenarios
n.sim <- 10 ^ 5
# correlation between Z1 and Z2
r <- 0.5
# simulation of Z1  and Z2
# constructed as a combination of independent standard normals U1, U2
U1 <- rnorm(n.sim)
U2 <- rnorm(n.sim)
Z1 <- 100 + 40 * U1
Z2 <- 100 + 20 * (r * U1 + sqrt(1 - r ^ 2) * U2)
# simulation of Z3
Z3 <- rnorm(n.sim, 100, 20)
# portfolio loss Y
Y <- Z1 + Z2 + Z3
# data of baseline model
dat <- data.frame(Z1, Z2, Z3, Y)
```


Now we introduce a stress to our baseline model. For our first stress, we require that the mean of $Z_1$ is increased from 100 to 110. This is done using the `stress` function, which generates as output the **SWIM** object `str.mean`. This object stores the stressed model, i.e. the realisations of the model components and the scenario weights. In the function call, the argument `k = 1` indicates that the stress is applied on the first column of `dat`, that is, on the realisations of the random variable $Z_1$.

``` {r, example1_first_stress, echo = -3, warning = FALSE, message = FALSE}
library(SWIM)
str.mean <- stress(type = "mean", x = dat, k = 1, new_means = 110)
options(digits = 2)
summary(str.mean, base = TRUE)
```

The `summary` function, applied to the **SWIM** object `str.mean`, shows how the distributional characteristics of all random variables change from the baseline to the stressed model. In particular, we see that the mean of $Z_1$ changes to its required value, while the mean of $Y$ also increases. Furthermore there is a small impact on $Z_2$, due to its positive correlation to $Z_1$.



Beyond considering the standard statistics evaluated via the `summary` function, stressed probability distributions can be plotted. In Figure \@ref(fig:example1-cdfs-mean) we show the impact of the stress on the cumulative distribution functions (cdf) of $Z_1$ and $Y$. It is seen how the stressed cdfs are lower than the original (baseline) ones. Loosely speaking, this demonstrates that the stress has increased (in a stochastic sense) both random variables  $Z_1$ and $Y$. While the stress was on $Z_1$, the impact on the distribution of the portfolio $Y$ is clearly visible.

``` {r, example1-cdfs-mean, warning = FALSE, message = FALSE, fig.show='hold', out.width = '50%', fig.cap = "Baseline and stressed empirical distribution functions of model components  $Z_1$ (left) and $Y$ (right), subject to a stress on the mean of $Z_1$."}
# refer to variable of interest by name...
plot_cdf(str.mean, xCol = "Z1", base = TRUE)
# ... or column number
plot_cdf(str.mean, xCol = 4, base = TRUE)
```

The scenario weights, given their central role, can be extracted from a **SWIM** object. In Figure \@ref(fig:example1-weights-mean), the scenario weights from `str.mean` are plotted against realisations from $Z_1$ and $Y$ respectively. It is seen how the weights are increasing in the realisations from $Z_1$. This is a consequence of the weights' derivation via a stress on the model component $Z_1$. The increasingness shows that those scenarios for which $Z_1$ is largest are assigned a higher weight. The relation between scenario weights and $Y$ is still increasing (reflecting that high outcomes of $Y$ tend to receive higher weights), but no longer deterministic (showing that $Y$ is not completely driven by changes in $Z_1$).


``` {r, example1-weights-mean, warning = FALSE, message = FALSE, fig.show='hold', out.width = '50%',fig.cap = "Scenario weights against observations of model components  $Z_1$ (left) and $Y$ (right), subject to a stress on the mean of $Z_1$."}
# extract weights from stressed model
w.mean <- get_weights(str.mean)
plot(Z1[1:5000], w.mean[1:5000], pch = 20, xlab = "Z1", ylab = "scenario weights")
plot(Y[1:5000], w.mean[1:5000], pch = 20, xlab = "Y", ylab = "scenario weights")
```


The stress to the mean of $Z_1$ did not impact the volatility of either $Z_1$ or $Y$, as can be seen by the practically unchanged standard deviations in the output of `summary(str.mean)`. Thus, we introduce an alternative stress that keeps the mean of $Z_1$ fixed at 100, but increases its standard deviation from 40 to 50. This new stress is seen to impact the standard deviation of the portfolio loss $Y$.

``` {r, example1_second_stress, cache = TRUE, echo = -2, warning = FALSE, message = FALSE}
str.sd <- stress(type = "mean sd", x = dat, k = 1, new_means = 100, new_sd = 50)
options(digits = 2)
summary(str.sd, base = FALSE)
```

Furthermore, in Figure \@ref(fig:example1-cdfs-sd), we compare the baseline and stressed cdfs of $Z_1$ and $Y$, under the new stress on $Z_1$. The crossing of probability distribution reflects the increase in volatility.


``` {r, example1-cdfs-sd, cache = TRUE, warning = FALSE, message = FALSE, fig.show='hold', out.width = '50%', fig.cap = "Baseline and stressed empirical distribution functions of model components  $Z_1$ (left) and $Y$ (right), subject to a stress on the standard deviation of $Z_1$."}
plot_cdf(str.sd, xCol = "Z1", base = TRUE)
plot_cdf(str.sd, xCol = 4, base = TRUE)
```

The different way in which a stress on the standard deviation of $Z_1$ impacts on the model, compared to a stress on the mean, is reflected by the scenario weights. Figure \@ref(fig:example1-weights-sd) shows the pattern of the scenario weights and how, when stressing standard deviations, higher weight is placed on scenarios where $Z_1$ is extreme, either much lower or much higher than its mean of 100.

``` {r, example1-weights-sd, cache = TRUE, warning = FALSE, message = FALSE, fig.show='hold', out.width = '50%',fig.cap = "Scenario weights against observations of model components  $Z_1$ (left) and $Y$ (right), subject to a stress on the standard deviation of $Z_1$."}
w.sd <- get_weights(str.sd)
plot(Z1[1:5000], w.sd[1:5000], pch = 20, xlab = "Z1", ylab = "scenario weights")
plot(Y[1:5000], w.sd[1:5000], pch = 20, xlab = "Y", ylab = "scenario weights")
```


Finally we ought to note that not all stresses that one may wish to apply are feasible. Assume for example that we want to increase the mean of $Z_1$ from 100 to 300, which exceeds the maximum realisation of $Z_1$ in the baseline model. Then, clearly, no set of scenario weights can be found that produce a stress that yields the required mean for $Z_1$; consequently an error message is produced.



```{r, example1_third_stress, cache = FALSE, error=TRUE, linewidth = 80}
stress(type = "mean", x = dat, k = 1, new_means = 300)
max(Z1)
```

