
# Simulation study {#Sec:CreditModel}

## The credit risk portfolio 

```{r, loading-packages, cache = FALSE, include = FALSE}
  library(SWIM)
  library(copula)
  options(digits = 3)
```


```{r, CM-data, cache = TRUE, include = FALSE}
#simulation of the credit model data (this will be removed)
  set.seed(1)
  nsim <- 100000
  
  #data
  m1 <- 2500 # counterparties tranche A
  m2 <- 5000 # counterparties tranche B
  m3 <- 2500 # counterparties tranche C
  
  p1 <- 0.0004 # prob of default 
  rho1 <- 0.0004 # correlation within the tranche

  p2 <- 0.0097 
  rho2 <- 0.0044
  
  p3 <- 0.0503  
  rho3 <- 0.01328
  
  #exposures
  e1 <- 80
  e2 <- 25 
  e3 <- 10 
  
  #loss given default
  LGD1 <- 0.25
  LGD2 <- 0.375
  LGD3 <- 0.5
  
  # beta-binomial model with copula 
  
  # beta parameters: matching tranches default probabilities and correlation
  alpha1 <- p1 * (1 / rho1 - 1)
  beta1 <- alpha1 * (1 / p1 - 1)
  
  alpha2 <- p2 * (1 / rho2 - 1)
  beta2 <- alpha2 * (1 / p2 - 1)
  
  alpha3 <- p3 * (1 / rho3 - 1)
  beta3 <- alpha3 * (1 / p3 - 1)
  
  # correlations between tranches
  cor12 <- 0.3
  cor13 <- 0.1
  cor23 <- 0.4
  
  # Gaussian copula structure
  myCop <- normalCopula(param = c(cor12, cor13, cor23), dim = 3, dispstr = "un")
  
  # define multivariate beta with given copula
  myMvd <- mvdc(copula = myCop,
                margins = c("beta", "beta", "beta"),
                paramMargins = list(list(alpha1, beta1),
                                    list(alpha2, beta2),
                                    list(alpha3, beta3)))
  
  # simulation from the chosen copula
  H <- rMvdc(nsim, myMvd)
  
  #simulate number of default per tranches (binomial distributions)
  M1 <- rbinom(n = nsim, size = m1, prob = H[, 1])
  M2 <- rbinom(n = nsim, size = m2, prob = H[, 2])
  M3 <- rbinom(n = nsim, size = m3, prob = H[, 3])
  
  #total loss per tranches
  L1 <- M1 * e1 * LGD1
  L2 <- M2 * e2 * LGD2
  L3 <- M3 * e3 * LGD3
  
  #global loss
  L <- L1 + L2 + L3
  
  #DB for SWIM
  credit_data <- cbind(L, L1, L2, L3, H)
  colnames(credit_data) <- c("L", "L1", "L2", "L3", "H1", "H2", "H3")
```



The credit model in this section is a conditionally binomial credit model and we refer to the Appendix \@ref(AppendixCM) for details and the generation of the simulated data. Of interest ist the total aggregate portfolio loss $L = L_1 + L_2 + L_3$, where $L_1, L_2, L_3$ are homogeneous subportfolios on comparable scale. The data set contains 100,000 simulations of the portfolio $L$, the sub-portfolios $L_1, L_2, L_3$ as well as the (conditional) default probability of each subportfolio $H_1, H_2, H_3$. A snipped of data set looks as follows:

```{r, CM-data-head, echo = FALSE, cache = TRUE}
options(digits = 3)
head(credit_data)
``` 



## Stressing the aggregate portfolio loss 

In this section, we study the effect of stresses on (the tail of) the aggregate portfolio on the three sub-portfolios. First, we stress the $VaR_{0.9}$ of the total loss of the aggregate portfolio by $20\%$. For this we use the `stress` function with the argument `type = "VaR"`. The input parameter `x` is the simulated data, `k` corresponds to the name of the row of `x` on which the stress is applied to, `alpha` determines the level of the stresses VaR and `q_ratio` the percentage increase.

```{r, CM-stress-VaR, cache = TRUE}
stress.credit <- stress(type = "VaR", x = credit_data, k = "L", 
                        alpha = 0.9, q_ratio = 1.2)
```

Second, we consider, additionally to the $20\%$ increase in $\text{VaR}_{0.9}$, a $30\%$ increase in $\text{ES}_{0.9}$ of the aggregate portfolio $L$. Generating a stressed model, resulting from a simultaneous stress on the VaR and the ES can be acchieve using `type = "VaR ES"`. Note that both VaR and ES need be stressed at the same level `alpha = 0.9`. The additonal input parameter `s_ratio` determines the percentage increase in the ES. Instead of providing the percentage inceases in the VaR and ES, the `stress` function allows for the actual stressed values of the VaR and ES using the parameters `s` and `q` instead of `s_ratio` and `q_ratio`, respectively. 


```{r, CM-stress-VaR-ES, cache = TRUE}
stress.credit <- stress(type = "VaR ES", x = stress.credit, k = "L", 
                        alpha = 0.9, q_ratio = 1.2, s_ratio = 1.3)
```

Note, that as input `x` we used the above calucalted stressed model, resulting from a stress on the $VaR_{0.9}$. Using a stressed model as an input for the `stress` function is convenient for large data sets, as the `stress` function returns an object (`stress.credit`) that countains the original simualted data and the scenario weights. 

++++MAYBE CHANGE THE SECOND STRESS ACCORDING TO ANDREAS' SUGGESTION? IE LEAVE VAR UNCHANGED (CURRENTLY NOT POSSIBLE) AND STRESS ES ONLY?


## Analysing the stressed model

The `summary` function provides a statistical summary of the stressed models. Choosing  `base = TRUE`, compares the stressed models with the the simulated data - the baseline model.

```{r, CM-summary, echo = -1, cache = TRUE}
options(digits = 3)
summary(stress.credit, base = TRUE)
```

The information on individual stresses can be recovered through the `get_specs` function and the actual scenario weitghts using `get_weight`.

```{r, CM-specs, echo = -1, cache = TRUE}
options(digits = 3)
get_specs(stress.credit)
w <- get_weights(stress.credit)
colMeans(w)
```
+++HERE WE COULD USE THE "expected shortfall" FUNCTION (NOT AVAILABLE YET) TO CALCULATE THE EXPECTED SHORTFALL OF "stress.credit$stress1" 



## Visual comparison
The change of the distributions of the portfolio and subportfolios from the baseline to the stressed models can be visualised through `plot_hist` and `plot_cdf`. The following plot displays the empirical histogramm of the aggregate portfolio loss under the baseline and the two stressed models.

```{r, CM-histL, cache = TRUE}
plot_hist(object = stress.credit, xCol = "L", base = TRUE)
```

Both functions,`plot_hist` and `plot_cdf`, include the parameters `xCol` specifying the columns of the data and `wCol` determining the columns of the scenario weights. Thus, allowing to plot the impact of the stressed models on the subportfoios. The graphical functions `plot_hist` and `plot_cdf` functions return objects compatible with the package **ggplot2**. Thus, we can compare the histograms of the portfolio losses via the function `grid.arrange` (of the package **gridExtra**).

```{r, CM-plot1, cache = FALSE}
library(gridExtra)
pL1 <- plot_hist(object = stress.credit, xCol = 2, wCol = 1, base = TRUE)
pL2 <- plot_hist(object = stress.credit, xCol = 3, wCol = 1, base = TRUE)
pL3 <- plot_hist(object = stress.credit, xCol = 4, wCol = 1, base = TRUE)
class(pL1)
grid.arrange(pL1, pL2, pL3, ncol = 1, nrow = 3)
```

From the plots we observe, that the subportfolios $L_2$ and $L_3$ are significanlty affected by the stress, while the distribution of $L_1$ is almost unchanged. 


## Sensitivity measures

The impact of the stressed models on the model components can be quantified through sensitivity measures. The function `sensitivity` includes the *Kolmogorov*, the *Wasserstein* distance and the sensitivity measure *Gamma*, which can be specified through the optional parameter `type`. We refer to Section \@ref(Sec:analysis) for the definition. The Kolmogorov and the Wasserstein distance are useful to compare different stressed models, whereas the sensitivity measure Gamma ranks model components for one stressed model.


```{r, CM-sensitivity1, cache = TRUE}
sensitivity(object = stress.credit, xCol = c(2 : 7), wCol = 1, type = "Gamma")
```

Using the `sensitivity` function we can analyse whether the first and third tranches are able to exceed the riskiness of the second. This can be accomplished specifying, through the option `f`, a list of functions applicable to the columns `k` of the dataset. Finally, setting `xCol = NULL` allows to consider only the transformed data:

```{r, CM-sensitivity2, cache = TRUE}
sensitivity(object = stress.credit,type = "Gamma", f = list(function(x)x[1] + x[2]), 
            k = list(c(2,4)), xCol = NULL, wCol = 1)
```

The `importance_rank` function, having the same structure as the `sensitivity` function, return the ranks of the sensitivity measures. This function is particularly useful when there are several risk factors involved.

```{r, CM-rank, cache = TRUE}
importance_rank(object = stress.credit, xCol = c(2 : 7), wCol = 1, type = "Gamma")  
```

It transpires that subportfolios $2$ and $3$ are, in this order, most responsible for the stress in the global loss. Also, most of the sensitivity seems to be due to the systematic risk components $H_2$ and $H_3$. To confirm this, another stress resulting in the same $\text{VaR}_{90\%}(L)$, but controlling the distribution of $H_2$, can be imposed using the function `stress_moment`. More precisely, we impose that $E[H_2]$ and the $75\%$ quantile of $H_2$ are fixed as in the base model. 

```{r, CM-stress-fixed-VaR}
VaR.L <- quantile(x = credit_data[, "L"], prob = 0.9, type = 1)
q.H2 <- quantile(x = credit_data[, "H2"], prob = 0.75, type = 1)
str.var.credit2 <- stress_moment(x = credit_data,
                                f = list(function(x)1 * (x <= VaR.L * 1.2),
                                         function(x)x,
                                         function(x)1 * (x <= q.H2)),
                                m = c(0.9, mean(credit_data[, "H2"]), 0.75),
                                k = c(1, 6, 6))
# stress.credit <- stress_moment(x = stress.credit,
#                                f = list(function(x)1 * (x <= VaR.L * 1.2),
#                                         function(x)x,
#                                         function(x)1 * (x <= q.H2)),
#                                m = c(0.9, mean(credit_data[, "H2"]), 0.75), k = c(1, 6, 6))
summary(str.var.credit2)
# summary(stress.credit)
sensitivity(object = str.var.credit2, xCol = c(2 : 7), type = "Gamma")
# sensitivity(object = stress.credit, xCol = c(2 : 7), type = "Gamma")
```

+++THIS SHOULD BE APPENDED TO "stress.credit" WHEN "stress_moment" IS FIXED
It is then clear that systematic risk prevails on binomial (event) risk. 

The `stress_moment` function is flexible and allows different type of stresses to be imposed on a model. The following example forces a $50\%$ increase in correlation between the losses in the second and third portfolios, while keeping the means und standard deviations unchanged.

```{r, CM-stress-moment, cache = TRUE}
m.L2 <- mean(credit_data[, "L2"])
m.L3 <- mean(credit_data[, "L3"])
m2.L2 <- mean(credit_data[, "L2"] ^ 2)
m2.L3 <- mean(credit_data[, "L3"] ^ 2)
cov.L2.L3 <- cov(credit_data[, "L2"], credit_data[, "L3"])
# str.var.credit2 <- stress_moment(x = credit_data,
#                                 f = list(function(x)x,
#                                          function(x)x,
#                                          function(x)x ^ 2,
#                                          function(x)x ^ 2,
#                                          function(x)x[1] * x[2] - m.L2 * m.L3),
#                                 k = list(3, 4, 3, 4, c(3, 4)),
#                                 m = c(m.L2, m.L3, m2.L2, m2.L3, cov.L2.L3 * 1.5)
```
+++CURRENTLY DOES NOT RUN - NEEDS TO BE FIXED OR REPLACED


+++FINAL COMMENTS?
